---
title: "Genomes in BSstring packages - Caenorhabditis elegans"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Caenorhabditis elegans reference genome structure

BSgenome.Celegans.UCSC.ce11 contains 7 Biostrings.


```{r, include=FALSE}
library(BSgenome.Celegans.UCSC.ce11)
eleg <- BSgenome.Celegans.UCSC.ce11
```

```{r, eval=FALSE}
library(BSgenome.Celegans.UCSC.ce11)
eleg <- BSgenome.Celegans.UCSC.ce11
```

Sequences have different lengh, with median 15279421
```{r}
seqlengths(eleg)
median(seqlengths(eleg))
```

After visualization is visible that Mitochondrial genome is significantly smaller. 
```{r, echo=FALSE}
plot(seqlengths(eleg), log = "y", ylab="log scale", main="Length of sequences"  )
```
 Let's look into first chromosome:
 
```{r}
eleg$chrI

```

To check what is the % of AT or GC in chr2L there is a fucntion: 

```{r, echo=FALSE}
af <- alphabetFrequency(eleg$chrI)
cat("percentage of AT:  ", as.character((af["A"] + af["T"])/nchar(eleg$chrI)))
cat("percentage of GC:  ", as.character((af["G"] + af["C"])/nchar(eleg$chrI)))

```

Lets check how many "TTAGGC" patter is in C.elegans
How many "TTAGGG" sequences are in the genome in reverse and froward? 

```{r}
dna_seq = DNAString("TTAGGC")
tel <- vmatchPattern(dna_seq, eleg)
length(tel)
```

What is the distribution of this pattern depending on the chromosome? According to literature repeated sequence in C. elegans is TTAGGC (http://telomerase.asu.edu/sequences_telomere.html)

```{r, echo=FALSE}
par(mfrow=c(3, 2))
empty <- seq(1, 6)
n=6
for (i in  1:n){
chr <- seqnames(eleg)[1:6][empty[i]]
hist(start(ranges(tel[seqnames(tel)==chr])),main = chr,breaks = 50,xlab = "Hits of TTAGGC")
} 
```

Mitochondrial DNA does not contain telomer patern at the ends. Why? Does mitochondrial DNA contain ends of chromosome? No, because is circullar.

```{r, echo=FALSE}
par(mfrow=c(1,1))
empty <- seq(1, 1)
n=1
for (i in  1:n){
chr <- seqnames(eleg)[7][empty[i]]
hist(start(ranges(tel[seqnames(tel)==chr])),main = chr,breaks = 50,xlab = "Hits of TTAGGC")
} 
```

And how about TTAGGC repeated 10 times in order? 


```{r}

dna_seq = DNAString("TTAGGC")
tel <- vmatchPattern(rep(dna_seq, 10), eleg)
length(tel)
```

What is the distribution of this pattern depending on the chromosome? According to literature repeated sequence in C. elegans is TTAGGC

```{r, echo=FALSE}
par(mfrow=c(3, 2))
empty <- seq(1, 6)
n=6
for (i in  1:n){
chr <- seqnames(eleg)[1:6][empty[i]]
hist(start(ranges(tel[seqnames(tel)==chr])),main = chr,breaks = 50,xlab = "Hits of TTAGGC")
} 
```



